---
title: "Testing palettes for color vision deficiency"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Testing palettes for color vision deficiency}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# knitr::opts_chunk$set(
#   collapse = TRUE,
#   comment = "#>"
# )
```

```{r setup, include = FALSE}
library(nmfspalette)
library(dichromat)
library(recolorize)
library(ggplot2)
```

# Background

We are testing that the existing NMFS color palettes are usable to someone with color vision deficiency (C.V.D.).

We'll use an R package called [`dichromat`](https://cran.r-project.org/web/packages/dichromat/index.html) to approximate how these palettes might be perceived by someone with any of the most common forms of C.V.D.: protanopia, deutanopia, and tritanopia.

This analysis is based on [this tutorial by Hannah Weller](https://hiweller.rbind.io/post/using-the-dichromat-package-to-check-if-your-plot-is-colorblind-friendly/).

## Install packages
```{r eval=FALSE}
install.packages("dichromat")
install.packages("recolorize")
install.packages("gridExtra")
```

## Load libraries
```{r}
library(dichromat)
library(recolorize)
library(gridExtra)
```

## Create a function that shows perceptions of palettes
```{r}
show_cvd_pals <- function(palette){
  # Convert palette into three palettes approximating how someone with each type of C.V.D. would perceive the original palette

  ## protanopia
  protan <- dichromat::dichromat(palette, type = "protan")
  ## deutanopia
  deutan <- dichromat::dichromat(palette, type = "deutan")
  ## tritanopia
  tritan <- dichromat::dichromat(palette, type = "tritan")
  
  # Plot the palettes and compare
  layout(matrix(1:4, nrow = 4))
  par(mar = rep(1, 4))
  
  recolorize::plotColorPalette(palette, main = "Trichromacy ('normal color vision')")
  recolorize::plotColorPalette(protan, main = "Protanopia")
  recolorize::plotColorPalette(deutan, main = "Deutanopia")
  recolorize::plotColorPalette(tritan, main = "Tritanopia")
}
```

# Test main palettes
```{r fig.align='center'}
# import main palettes from nmfs_cols.R file
core_nmfs_pals <- list(
  `oceans` = nmfs_cols("pale_sea_blue", "light_sea_blue", "processblue", "reflexblue", "national", "o50pblack"),
  `waves` = nmfs_cols("pale_teal", "bright_teal", "westcoast", "dark_teal_green"),
  `seagrass` = nmfs_cols("bright_seagrass", "ltgreen", "southeast", "dark_seagrass"),
  `urchin` = nmfs_cols("bright_urchin", "vivid_urchin", "midatlantic", "dark_urchin"),
  `crustacean` = nmfs_cols("bright_crustacean", "ltorange", "alaska", "dark_crustacean"),
  `coral` = nmfs_cols("bright_coral", "vivid_coral", "pacificislands", "dark_coral"),
  `regional` = nmfs_cols("national", "westcoast", "southeast", "midatlantic", "alaska", "pacificislands")
)

# Show comparisons for each palette
for (i in 1:length(core_nmfs_pals)){
  print(paste0(names(core_nmfs_pals)[i], " palette"))
  show_cvd_pals(core_nmfs_pals[[i]])
}
```

**Conclusion**: For someone with C.V.D., the regional palette could be challenging to discern. In particular, colors look very similar when viewed with the following perceptions: protanopia (3, 5); deutanopia (3, 5, 6); tritanopia (2, 3, 4).

# Test alternative regional color palette
## Create alternative palettes
```{r}
# Option 1: using existing NMFS colors
regions_alt1 <- c("#002364", # "national"
                  "#00797F", # "westcoast"
                  "#76BC21", # "ltgreen"
                  "#5761C0", # "midatlantic"
                  "#DB6015", # "alaska"
                  "#B71300"  # "pacificislands"
                  )

# Option 2: using some colors outside of the NMFS color set
regions_alt2 <- c("#016080", # new color
                  "#57aeb2", # new color
                  "#26DC61", # new color
                  "#8851D0", # new color
                  "#DB6015", # "alaska"
                  "#B71300"  # "pacificislands"
                  )

alt_regional_pals <- list(`Alternative regional palette 1` = regions_alt1,
                          `Alternative regional palette 2` = regions_alt2)

```

## Show alternative palettes
```{r fig.align='center'}
for (i in 1:length(alt_regional_pals)){
  print(names(alt_regional_pals)[i])
  show_cvd_pals(alt_regional_pals[[i]])
}
```

## Show alternative palettes in figures
```{r message=FALSE, warning=FALSE, fig.width=7, fig.height=5}
show_cvd_plots <- function(pal,
                           pal_name){
  
  test_data <- data.frame(
  palette_color = c("1", "2", "3", "4", "5", "6",
                    "1", "2", "3", "4", "5", "6"),
  y = c(1,2,3,4,5,6,
        2,3,4,5,6,7)
  )
  
  type <- "Trichromacy ('normal color vision')"
  orig <- ggplot2::ggplot(data = test_data,
          aes(x = palette_color,
              y = y,
              color = palette_color)) + 
    geom_point(size = 4,
               shape = 19) +
    geom_line(size = 1) +
    theme_light() +
    theme(legend.position = "none") +
    labs(subtitle = type,
         x = "Palette color") +
    scale_color_manual(values = pal)
  orig
  
  type <- "Protanopia"
  protan <- orig +
    labs(subtitle = type) +
    scale_color_manual(values = dichromat::dichromat(pal, type = "protan")) 
  protan
  
  type <- "Deutanopia"
  deutan <- orig +
    labs(subtitle = type) +
    scale_color_manual(values = dichromat::dichromat(pal, type = "deutan"))
  deutan
  
  type <- "Tritanopia"
  tritan <- orig +
    labs(subtitle = type) +
    scale_color_manual(values = dichromat::dichromat(pal, type = "tritan"))
  tritan
  
  four_plots <- gridExtra::grid.arrange(orig, protan, deutan, tritan,
                          top = pal_name)
  
  return(four_plots)
}

for (i in 1:length(alt_regional_pals)){
  show_cvd_plots(pal = alt_regional_pals[[i]],
                 pal_name = names(alt_regional_pals)[i])
}
```

**Conclusion**: For someone with C.V.D., the alternative palette 1 shows colors 2 & 4 as very similar for tritanopia. Alternative palette 2 provides colors that seem distinguishable for all types of color perception, but multiple colors in this palette are outside the NMFS brand color set.

Both alternative regional palettes ("regional_alt1" and "regional_alt2", respectively) are available in the package.
