---
title: "NOAA Themed Presentations"
author: "Christine Stawitz"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{NOAA Themed Presentations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
install.packages("remotes")
remotes::install_github("gadenbuie/xaringanthemer")
library(xaringanthemer)
# remotes::install_github("nmfs-fish-tools/nmfspalette")
library(nmfspalette)
```

## Theming presentations

The `docs/` folder of the package includes an `.Rmd` called `xaringan_themer.Rmd` that allows you to create a themed xaringan presentation. The themer uses the `nmfspalette` package to source the colors for text, background, and footnotes and then the `xaringanthemer` package to apply the formatting to slides without having to write CSS.

The `xaringanthemer` package is needed to use this template code. Custom theme settings can be modified using the `style_xaringan()` function. In general, the arguments to `style_xaringan` follow [CSS property](https://www.w3schools.com/cssref/) naming style and syntax. You don't need to know CSS to use it, but it can be helpful. The `style_xaringan()` function generates a .css file called `xaringan_themer.css` that you will need to add into the 

First, we will look at the arguments to change colors of text and backgrounds, which all end in `_color`. The following code changes the background, text, footnote, link, and code highlight color to match NOAA Fisheries theming.



# Color theming

```{r theming, exercise = TRUE}
style_xaringan(
  title_slide_background_color = unname(nmfs_cols("darkblue")),
  title_slide_text_color = unname(nmfs_cols("supltgray")),
  background_color = unname(nmfs_cols("white")),
  text_color = unname(nmfs_cols("darkblue")),
  header_color = unname(nmfs_cols("darkblue")),
  inverse_background_color = unname(nmfs_cols("processblue")),
  inverse_text_color = unname(nmfs_cols("supltgray")),
  footnote_color = unname(nmfs_cols("darkblue")),
  link_color = unname(nmfs_cols("medteal")),
  text_slide_number_color = unname(nmfs_cols("lightteal")),
  code_highlight_color = unname(nmfs_cols("medteal"))
)
```

# Font theming

To change fonts, we use the arguments ending in `_font_google`. To change the font size, use the arguments ending `_font_size`. We also change the footnote position using the `footnote_position_bottom` argument. You can provide the font size in *em* or *px* units. Font arguments need to be wrapped in `google_font()` or `xaringanthemer_font_default` wrappers. If you plan to use Google fonts, search for names on the web to find fonts similar to your choice - searching for "Calibri Google font" yielded their equivalent: "Carlito".

```{r fonts, exercise = TRUE}
style_xaringan(
  header_font_google = google_font("Arimo"),
  text_font_google = google_font("Carlito", "300", "300i"),
  code_font_google = google_font("Source Code Pro"),
  footnote_font_size = "0.6em",
  footnote_position_bottom = "10px",
  text_slide_number_font_size = "0.6em"
)
```

# Background images

```{r background_images, exercise = TRUE}
# Find package directory
images_path <- system.file("static", package = "nmfspalette")

style_xaringan(
  background_image = file.path(images_path, "slideswoosh.PNG"),
  background_size = "cover",
  title_slide_background_image = file.path(images_path, "slideswooshver.png"),
  title_slide_background_size = "cover"
)
```

# Extra css
  
```{r extracss, exercise = TRUE}
style_xaringan(extra_css = list(".remark-slide-number" = list(
  "font-size" = "0.6em",
  "font-weight" = "bold",
  "margin" = "0px 840px -2px 0px"
)))
```

